A National Pipeline System as a Solution to Future Water Shortage in the United States

### 2013-02-04 paper.tex
Continue work on Analysis/Vis

Due to the scale and complexity of the task of water management in
the United States, we make several simplifying assumptions in order
to obtain a workable model.
\begin{itemize}
\item Counties are thought of as discrete entities, occupying points in
space located at their centroid. This assumption enables us to think
of counties as vertices (as in a graph). This assumption can be made
without losing too much accuracy for several reasons: The United States
is split into more than 3,000 counties, with an average area of less
than 3,300 square kilometers, and average length across of less than
60 kilometers. The level of inaccuracy is sufficiently low that it
is workable (in contrast, states for instance are far too large to
be accurately modeled as points).
\item De-salinization plants, reservoirs, and purification plants are all facilities that are built at a single location, and have the intended effect of providing more water to that location. For the purposes of our modelling, we can think of them as being essentially the same structure, with the proviso that they each can only be constructed in particular parts of the country. 
\item When considering the costs of pipelines, only the distance of pipeline
is considered. The landscape of a region may in practice affect the
length of pipeline required, or may result in additional cost per
length of pipeline, but we shall not model this, as we could not find
data on the difficulty of pipeline construction across different types
of terrain.
\item Pipelines are thought of as bi-directional. This assumption essentially
reflects the reality of pipelines. There are a few exceptional cases
where it is significantly more difficult to send water in one direction,
but these are rare enough that they can be disregarded.
\item Pipelines have variable rates of water transmission, depending on
factors such as size. In theory it would be possible to save money on pipelines by constructing smaller pipelines in areas where water rarely needs to be transported at high rates. 
However, the amount of water that needs to be transported through any given pipe is highly variable (ie. after a sudden intense rainfall). Thus, pipelines for long-distance water transmissions are relatively stable in cost. Our estimate below for pipeline cost is a typical figure.
\item Some pipelines exist already, but we do not consider them in our model. This is partially because they are likely to have minimal impact because there are not many of them, and partially because already-existing pipelines are poorly documented, so there is no way to efficiently collect and parse such data
\end{itemize}

@@

The cost of a de-salinization facility in 2013 is approximately \$14
per gallon per day\cite{key-12}. This figure is approximately the same as for similar kinds of facilities (reservoirs, dams).

@@

Edges in the Minimal Spanning Tree in Figure \ref{fig:graph} are weighted inversely proportional to cost and directly proportional to the populations at each of the endpoints, and sorted. The resulting sorted table gives a full list of all $3108$ pipelines sorted by this utility metric. The final (highest priority) rows are shown in Table \ref{table:priority}, with FIPS1 and FIPS2 representing the FIPS codes of the ends of each pipe.

\begin{table}[h]\label{table:priority}
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
FIPS1 & FIPS 2 & Priority \tabularnewline
\hline
48201 & 48339 & 10346.3 \tabularnewline
\hline
24005 & 24510 & 10675.4 \tabularnewline
\hline
6001  &6013 & 11019.8 \tabularnewline
\hline
34003 & 34017 & 11190.7 \tabularnewline
\hline
51153 & 51685 & 12104.5 \tabularnewline
\hline
11001 & 51013 & 12113.6 \tabularnewline
\hline
48029 & 48091 & 13787.1 \tabularnewline
\hline
34017 & 36061 & 16584.3 \tabularnewline
\hline
51059 & 51600 & 18528.8 \tabularnewline
\hline
51153 & 51683 & 22134.2 \tabularnewline
\hline
6037 & 6059  &23742.4 \tabularnewline
\hline
36047 & 36061  &27446.1 \tabularnewline
\hline
36059 & 36081  &29256.5 \tabularnewline
\hline
36005 & 36061 & 32595 \tabularnewline
\hline
36047 & 36081 & 33517.6 \tabularnewline
\hline
17031 & 17043 & 41637.4 \tabularnewline
\hline

@@

\section{Potential Weaknesses}

Our proposed pipeline network can play a huge role in managing the distribution of freshwater over the United States. However, it is not a perfect solution to water management problems, and has some potential flaws:
\begin{itemize}
\item Before completion, there are holes in the network: Before every pipeline has been constructed, 
there will be some pairs of counties that are not connected by pipelines. In this case, the network is still 
useful, but is not quite as powerful as a complete network would be.
\item Large scale droughts can be an issue: if the entirety of the one region (say, the west coast) is in 
drought, while (say) the east coast has a water surplus, there may be bottlenecks in the middle of the country 
as large amounts of water are transferred. This could potentially be an issue, but in reality, this kind of situation 
is extremely rare, partially because the coasts of the United States do not tend to be where droughts occur.
\item Some nearby counties are only connected by long and convoluted paths: This is not as much of an issue as 
one might imagine, because generally droughts happen slowly, and hence it is acceptable for water to take a long time to 
reach its destination. However, if this is seen as a major issue, a small number of additional pipes can be added on a case-by-case basis.
\end{itemize}

However, none of these is a fatal flaw in the idea of constructing a pipeline network.

>>>>

Due to the scale and complexity of the task of water management in
the United States, we make several simplifying assumptions in order
to obtain a workable model.
\begin{itemize}
\item Counties are thought of as discrete entities, occupying points in
space located at their centroid. This assumption enables us to think
of counties as vertices (as in a graph). This assumption can be made
without losing too much accuracy for several reasons: The United States
is split into more than 3,000 counties, with an average area of less
than 3,300 square kilometers, and average length across of less than
60 kilometers. The level of inaccuracy is sufficiently low that it
is workable (in contrast, states for instance are far too large to
be accurately modeled as points).
\item De-salinization plants, reservoirs, and purification plants are all facilities that are built at a single location, and have the intended effect of providing more water to that location. For the purposes of our modelling, we think of them as being essentially the same kind of thing.
\item When considering the costs of pipelines, only the distance of pipeline
is considered. The landscape of a region may in practice affect the
length of pipeline required, or may result in additional cost per
length of pipeline, but we shall not model this, as we could not find
data on the difficulty of pipeline construction across different types
of terrain.
\item Pipelines are thought of as bi-directional. This assumption essentially
reflects the reality of pipelines. There are a few exceptional cases
where it is significantly more difficult to send water in one direction,
but these are rare enough that they can be disregarded.
\item Pipelines have variable rates of water transmission, depending on
factors such as size. In theory it would be possible to save money on pipelines by constructing smaller pipelines in areas where water rarely needs to be transported at high rates. 
However, the amount of water that needs to be transported through any given pipe is highly variable (ie. after a sudden intense rainfall). Thus, pipelines for long-distance water transmissions are relatively stable in cost. Our estimate below for pipeline cost is a typical figure.
\item Some pipelines exist already, but we do not consider them in our model. This is partially because they are likely to have minimal impact because there are not many of them, and partially because already-existing pipelines are poorly documented, so there is no way to efficiently collect and parse such data
\end{itemize}

@@

The cost of a de-salinization facility in 2013 is approximately \$14
per gallon per day\cite{key-12}.

@@

Edges in the Minimal Spanning Tree in Figure \ref{fig:graph} are weighted inversely proportional to cost and directly proportional to the populations at each of the endpoints, and sorted. The resulting sorted table gives a full list of all $3108$ pipelines sorted by this utility metric. The final (highest priority) rows are shown in Table \ref{table:priority}.

\begin{table}[h]\label{table:priority}
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
FIPS 1 & FIPS 2 & Priority \tabularnewline
\hline
48201 & 48339 & 10346.3 \tabularnewline
\hline
24005 & 24510 & 10675.4 \tabularnewline
\hline
6001  &6013 & 11019.8 \tabularnewline
\hline
34003 & 34017 & 11190.7 \tabularnewline
\hline
51153 & 51685 & 12104.5 \tabularnewline
\hline
11001 & 51013 & 12113.6 \tabularnewline
\hline
48029 & 48091 & 13787.1 \tabularnewline
\hline
34017 & 36061 & 16584.3 \tabularnewline
\hline
51059 & 51600 & 18528.8 \tabularnewline
\hline
51153 & 51683 & 22134.2 \tabularnewline
\hline
6037 & 6059  &23742.4 \tabularnewline
\hline
36047 & 36061  &27446.1 \tabularnewline
\hline
36059 & 36081  &29256.5 \tabularnewline
\hline
36005 & 36061 & 32595 \tabularnewline
\hline
36047 & 36081 & 33517.6 \tabularnewline
\hline
17031 & 17043 & 41637.4 \tabularnewline
\hline

@@
